---
title: "Task 8 - Functions"
output:
  md_document:
    variant: markdown_github
---

# Functions and error handling: #

## Write a function that fulfills the following criteria: ##

*It should be tidyverse compatible (i.e., the first argument must be a data frame)
*It should add two arbitrary columns of the data frame (specified by the user) and putthem in a new column of that data frame, with the name of the new column specified by the user
*It should throw an informative warning if any invalid arguments are provided. Invalid arguments might include:
**The first argument is not a data frame
**Less than two valid columns are specified to add (e.g., one or both of the column names isn't in the supplied data frame)
**The columns specified are not numeric, and therefore can't be added - use tryCatch() for this
*If the columns to add aren't valid but the new column name is, the function should create a column of NA values


```{r}
library(tidyverse)

#Function to add values in columns
add_column <- function(data_frame, col.one, col.two, sum.col) {
  
  if(class(data_frame) != "data.frame") {
    warning("Data frame not supplied") #warning for if data_frame argument is not a data frame
  }
  
  if(!(col.one %in% colnames(data_frame)) | !(col.two %in% colnames(data_frame))) {
    warning("One of your columns in not in the supplied data frame") #warning for if either columns to add are not actually columns in given data frame
  }
  
  tryCatch(data_frame$sum.col <- (data_frame[, col.one] + data_frame[, col.two]),  #runs function and returns warnings if column values are non-numeric
           warning=function(w){
             message("One or more values to add is non-numeric.")
             return(NA)
  })
  
  return(data_frame)
}

#Code lines to generate warnings and errors; using mtcars as an example

mtcars_adj <- mtcars 
mtcars_adj <- add_column(mtcars, "mpg", "cyl", sum) 
mtcars_adj$names <- rownames(mtcars_adj)
new_mtcars <- add_column(mtcars_adj, "mpg", "cyl", sum)
#new_mtcars <-add_column(mtcars_adj, "names", "mpg", sum) ***throws an error as "names" has no numeric values
```

# Loop and performance metric tasks


*Write a function named that uses a for loop to calculate the sum of the elements of a vector, which is passed as an argument (i.e., it should do the same thing that sum() does with vectors). your_fun(1:10^4) should return 50005000.
*Use the microbenchmark::microbenchmark function to compare the performace of your function to that of sum in adding up the elements of the vector 1:10^4. The benchmarking code should look something like:

```{r}

library(microbenchmark)

new.vec <- c(2,4,6,8,10)

#add.vec.columns <- add_column(new.vec, "mpg", "cyl", sum) ***throws an error as a data frame is not supplied, and correct columns not present

col_adder <- function(my.vec) {
  sum <- 0
  for(i in my.vec) {
    sum <- sum + i
  }
  return(sum)
}

#Test the function 
test.vec <- 1:10^4 

#Returns sum of function created by me
col_adder(test.vec) 

#Returns function using R nested function
sum(test.vec) 

#Tests the efficiency of my function
microbenchmark( 
  col_adder(test.vec),
  sum(test.vec)
)
```

