---
title: "Task 7 : Dplyr"
output: 
    md_document:
        variant: markdown_github
---

## Using the nycflights13::weather:
# Determine whether there are any clear outliers in wind speed (wind_speed) that should be rejected. If so, filter those bad point(s) and proceed.

```{r}
library(dplyr)
library(tidyverse)
library(nycflights13)

# Makes a data frame table with nycflights weather
nyc.weather <- nycflights13::weather

# Determine outliers that are contained in the data visually
nycwind_plot <- ggplot(nyc.weather, aes(x=time_hour, y=wind_speed)) + geom_point()
print(nycwind_plot)                       

# Filter out the outliers that are artifacts from data collection
windspeed_filtered <- filter(nycflights13::weather, wind_speed < 200)
```

## What direction has the highest median speed at each airport? Make a table and a plot of median wind speed by direction, for each airport.

```{r}
### Find direction with the highest median speed at each airport 

# Creates a dataframe that groups each airport by their wind speed
windspeed_dplyr <- nyc.weather %>%
  group_by(wind_speed) %>%
  filter(wind_speed < 200) %>%
  print(windspeed_dplyr)

# Creates a dataframe that contains each airport and their median speed for each direction
med_windspeed = windspeed_dplyr %>%
  group_by(wind_dir, origin) %>%
  summarise(wind_spd_med = median(wind_speed)) %>%
  group_by(origin) %>%
  filter(wind_spd_med == max(wind_spd_med))

##### Makes tables and plots of median wind speed by direction, for each airport.
## Respective plots show the importance in choosing the right type of plot to covey your data.


EWR = med_windspeed %>%
  filter(origin == 'EWR')

ggplot(EWR, aes(wind_spd_med, wind_dir)) +
  geom_point() +
  xlab('Wind Speed') +
  ylab('Wind Direction')

JFK = med_windspeed %>%
  filter(origin == "JFK")

ggplot(JFK, aes(wind_spd_med, wind_dir)) +
  geom_line() +
  xlab('Wind Speed') +
  ylab('Wind Direction')

LGA = med_windspeed %>%
  filter(origin == "LGA")

ggplot(JFK, aes(wind_spd_med, wind_dir)) +
  geom_boxplot() +
  xlab('Wind Speed') +
  ylab('Wind Direction')
```

## Using nycflights13::flights and nycflights13::airlines:
# Make a table with two columns: airline name (not carrier code) and median distance flown from JFK airport. The table should be arranged in order of decreasing mean flight distance. Hint: use a _join function to join flights and airlines.
# Make a wide-format data frame that displays the number of flights that leave Newark ("EWR") airport each month, from each airline

```{r}
# Makes a table that joins flights and airlines
nyc_flights_air <- inner_join(nycflights13::flights, airlines)

# Filters the nyc_flights_air table to include only the airline and mean distance ranked from highest to lowest
JFK_med_dist = nyc_flights_air %>%
  filter(origin == "JFK") %>%
  group_by(name) %>%
  summarise(med_dist_JFK = median(distance)) %>%
  arrange(desc(med_dist_JFK))
```

# Make a wide-format data frame that displays the number of flights that leave Newark ("EWR") airport each month, from each airline

```{r}
# Creates a wide format data frame that contains the number of flights that leave EWR each month
wide.frame <- nyc_flights_air %>%
  group_by(month, name) %>%
  summarise(month.airline = length(month)) %>%
  spread(name, month.airline)
```

## Using the babynames dataset:
# Identify the ten most common male and female names in 2014. Make a plot of their frequency (prop) since 1880. (This may require two separate piped statements).

```{r}
library(babynames)

# Creates a table that contains baby names dataset
baby_names <- babynames::babynames
 
# Filters out top 10 names in 2014
## Jacob is number 4!
topnames = baby_names %>%
  filter(year == 2014) %>%
  group_by(sex) %>%
  top_n(10,n) %>%
  select(year, sex, name)

# Filters out the top 10 girl names in 2014
topgirl = baby_names %>%
  filter(year == 2014) %>%
  filter(sex == 'F') %>%
  top_n(10,n)

# Filters out the top 10 boy names in 2014 
topboy = baby_names %>%
  filter(year == 2014) %>%
  filter(sex == 'M') %>%
  top_n(10,n)

# Creates a table of the top names for girls between 1880 to 2014 by their frequency
name_girl <- baby_names %>%
  filter(sex == 'F') %>%
  filter(name %in% topgirl$name) %>%
  select(year,sex,name,prop) %>%
  group_by(name)

# Creates a table of the top names for girls between 1880 to 2014 by their frequency
name_boy <- baby_names %>%
  filter(sex == 'M') %>%
  filter(name %in% topboy$name) %>%
  select(year,sex,name,prop) %>%
  group_by(name)

# Plot that show frequency of girl names since 1880
ggplot(name_girl, aes(year, prop, color = name))+
  geom_line()

# Generates a line plot that shows frequency of boy names since 1880
ggplot(name_boy, aes(year, prop, color = name))+
  geom_line()

# Creates a table of common female names in 1896
common_1896 <- baby_names %>%
  select(year, sex, name, n) %>%
  filter(sex == 'F') %>%
  filter(year == 1896)

# Creates a table of common female names in 1942
common_1942 <- baby_names %>%
  select(year, sex, name, n) %>%
  filter(sex == 'F') %>%
  filter(year == 1942)

# Creates a table of common female names in 2014
common_2014 <- baby_names %>%
  select(year, sex, name, n) %>%
  filter(sex == 'F') %>%
  filter(year == 2014)
```

#Make a single table of the 26th through 29th most common girls names in the year 1896, 1942, and 2016

```{r}
# Joins common names for females in 1896, 1942, and 2014
common_names_join = full_join(common_1896, common_1942)
common_names = full_join(common_names_join, common_2014)

# Generates a table with the most 26th through 29th most common names amongst girls in 1896, 1942, and 2014 
age26_29 = common_names %>%
  group_by(year) %>%
  top_n(29,n) %>%
  top_n(-4,n)
print(age26_29)
```

##Write task that involves some of the functions on the Data Wrangling Cheat Sheet and execute it.

```{r}
#### Write task that involves some of the functions on the Data Wrangling Cheat Sheet and execute it. 
## Determine how the frequency of girls to be named Tyler, which is my middle name, from 1880 to 2014.

# Creates a table of the frequency of girls named Tyler from 1880 to 2014
girl_Tyler = baby_names %>%
  filter(name == 'Tyler') %>%
  filter(sex == 'F')

# Generates a plot that visually shows the frequency of girls named Tyler
ggplot(girl_Tyler, aes(year, n)) +
  geom_line(stat = "identity") +
  xlab("year") +
  ylab("number of girls")
```

